---
title: "Data Tidying"
author: "Patricia DeRepentigny"
date: "1/16/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

```{r}
# Load necessary packages
library(dplyr)
library(tidyr)
```

If a function has the same name in two different packages, the one from the last loaded package will be the one working.

To call a function from a specfic package: `package_name::function_name(...)`

To remove all packages loaded: Session -> Restart R

# Data Cleaning

Read in data file

```{r, echo=F}
catch <- read.csv(file = "https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1")
head(catch)
```

Pipe operator: cmd + shift + M (%>%)

Steps covered here:  
* Remove marginal sum and notes columns  
* Move from wide to long format  

```{r, echo=F}
catch_long <- catch %>% 
  select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum) %>% # or select(-All, -notesRegCode)
  gather(key = "Species", value = "Catch", -Year, -Region)

head(catch_long)
```

Changed the name of the Catch column to Catch_thousands to better represents that catch numbers are in thousands of fish.

```{r, echo=F}
catch_cleaned <- catch_long %>% 
  rename(Catch_thousands = Catch) #%>% 
  #mutate(Catch = Catch_thousands * 1000)

head(catch_cleaned)
```

```{r}
test_catch <- as.integer(catch_cleaned$Catch_thousands)

i <- which(is.na(test_catch) == TRUE)

catch_cleaned[i, ]
```

# Split-Apply-Combine

# Joins
