---
title: "Data Tidying"
author: "Patricia DeRepentigny"
date: "1/16/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r}
# Load necessary packages
library(dplyr)
library(tidyr)
```

If a function has the same name in two different packages, the one from the last loaded package will be the one working.

To call a function from a specfic package: `package_name::function_name(...)`

To remove all packages loaded: Session -> Restart R

# Data Cleaning

Read in data file

```{r, echo=F}
catch <- read.csv(file = "https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1")
head(catch)
```

Pipe operator: cmd + shift + M (%>%)

Steps covered here:  
* Remove marginal sum and notes columns  
* Move from wide to long format  

```{r, echo=F}
catch_long <- catch %>% 
  select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum) %>% # or select(-All, -notesRegCode)
  gather(key = "Species", value = "Catch", -Year, -Region)

head(catch_long)
```

Steps covered next:  
* Changed the name of the Catch column to Catch_thousands to better represents that catch numbers are in thousands of fish  
* Erroneous value due to OCR issue - change I to 1  
* Create a new column for Catch in exact numbers  

```{r, echo=F}
catch_cleaned <- catch_long %>% 
  rename(Catch_thousands = Catch) %>%
  mutate(Catch_thousands = ifelse(Catch_thousands == "I", 1, Catch_thousands)) %>% # or 1, as.integer(Catch_thousands)
  mutate(Catch_thousands = as.integer(Catch_thousands)) %>% 
  mutate(Catch = Catch_thousands * 1000)

tail(catch_cleaned)
```

```{r, eval=F, echo=F}

# I used this code to find the bad value
test_catch <- as.integer(catch_cleaned$Catch_thousands)

i <- which(is.na(test_catch) == TRUE)

catch_cleaned[i, ]
```

# Split-Apply-Combine

Calculate the total catch by region

```{r, echo=F}
catch_total <- catch_cleaned %>% 
  group_by(Region, Year) %>% 
  summarise(Catch_Region_Year = sum(Catch), n_obs = n())

head(catch_total)
```

Matt's challenge

```{r, echo=F}
catch_matt <- catch_cleaned %>% 
  group_by(Year, Species) %>% 
  summarise(Catch_Mean = mean(Catch), Catch_sd = sd(Catch), n_obs = n())

head(catch_matt)
```

Filter for chinook salmon

```{r, echo=F}
catch_chinook <- catch_cleaned %>% 
  filter(Species == "Chinook" & Region == "SSE" & Year >= 1990) %>% 
  arrange(-Year)

head(catch_chinook)
```

# Joins

Read in region definitions files

```{r, echo=F}
region_defs <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1", method = "libcurl"),
                        stringsAsFactors = FALSE)

head(region_defs)
```

Clean up region definitions data

```{r, echo=F}
region_cleaned <- region_defs %>% 
  select(code, mgmtArea)

head(region_cleaned)
```

```{r, echo=F}
catch_joined <- left_join(catch_cleaned, region_cleaned,
                          by = c("Region" = "code"))

head(catch_joined)
```

## Spread

Make a wide dataframe using spread

```{r}
catch_wide <- catch_cleaned %>% 
  filter(Year > 1990) %>%
  select(-Catch_thousands) %>% 
  spread(key = Year, value = Catch)

head(catch_wide)
```

## Separate and Unite

```{r}
dates_df <- data.frame(date = c("5/24/1930",
                                "5/25/1930",
                                "5/26/1930",
                                "5/27/1930",
                                "5/28/1930"),
                       stringsAsFactors = FALSE)

dates_df
```

```{r}
dates_sep <- dates_df %>% 
  separate(col = date, into = c("month", "day", "year"), sep = "/", remove = FALSE)

dates_sep
```

```{r}
dates_unite <- dates_sep %>% 
  unite(date_iso, year, month, day, sep = "-")

dates_unite
```

